---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# speakfriend

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/BenjaminWolfe/speakfriend.svg?branch=master)](https://travis-ci.org/BenjaminWolfe/speakfriend)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/BenjaminWolfe/speakfriend?branch=master&svg=true)](https://ci.appveyor.com/project/BenjaminWolfe/speakfriend)
[![Codecov test coverage](https://codecov.io/gh/BenjaminWolfe/speakfriend/branch/master/graph/badge.svg)](https://codecov.io/gh/BenjaminWolfe/speakfriend?branch=master)
<!-- badges: end -->

`speakfriend` is designed for R champions and enthusiasts
who want to make analysis easy at their school or place of work,
without perhaps having a comprehensive solution such as Rstudio Server (yet!).

[RStudio][1]'s [keyring][2] package provides secure password management for R,
as [Sharon Machlis][3] outlines in [this excellent article][4].

`speakfriend` adds 4 elements targeted to the beginner:

1. __Descriptive prompts__.
   `keyring` shows the same featureless prompt no matter what's being asked.
   This can be confusing for new users who need to know clearly
   whether R needs their _master password_ or an individual _system password_,
   and why.
1. __Supportive default behaviors__.
   For example, if the user tries to create a key
   but doesn't yet have a keyring to store it,
   `speakfriend` will offer to create one first, and explain what it's doing.
1. __Instructive messages__.
   `speakfriend` strives to give the user as much information as is helpful.
   For example if an action that requires interactivity
   is being run outside an interactive session (e.g. unlocking the keyring),
   `speakfriend` will fail and tell the user
   exactly what function they need to run.
1. __Streamlined functionality__.
   `speakfriend` uses just one keyring and stores just one password per system.
   This is all many users need.
   `keyring` allows more options, outlined below under _Nomenclature_.

`keyring` is a powerful and flexible package.
`speakfriend` seeks to address tasks important to developing
an active R community.
These are what [Emily Riederer][5] calls [first-mile tasks][6],
tasks you should make as simple as possible, as early as possible,
to empower and engage your community.

## Installation

`speakfriend` is not yet available on [CRAN][7].
You can install it from [GitHub][8] with:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("BenjaminWolfe/speakfriend")
```

## Example

`speakfriend` has just 11 exported functions:

```{r example, eval=FALSE}
library(speakfriend)

create_keyring()        # creates keyring, leaves it in unlocked state
has_keyring()           # checks for existence of keyring
lock_keyring()          # locks keyring, does not require master password
keyring_locked()        # checks if keyring is locked
unlock_keyring()        # prompts for master password, unlocks keyring

list_keys()             # lists keys
set_key("system_name")  # prompts for individual password, sets key
has_key("system_name")  # checks for existence of key for a given system
get_key("system_name")  # returns individual password
drop_key("system_name") # forgets individual password

drop_keyring()          # e.g. if you forgot your master password
```

## Package Name Inspiration

`speakfriend` is named after a scene
[in J.R.R. Tolkien's The Fellowship of the Ring][9].
The party comes upon an enormous stone door with an elven inscription:

> The Doors of Durin, Lord of Moria. Speak, friend, and enter.

They tried several passwords, spells, and gestures,
until finally it occurred to Gandalf that it was a simple riddle.
They needed only to say the elven word for "friend" and the door would open.

`speakfriend` aims for a similar level of simplicity;
one might say you type `speakfriend`, and `enter`.
And as the Doors of Durin once [allowed unfettered trade][10]
between the words of the Dwarves and the Elves,
hopefully this package will help you develop a thriving community of R users.

I considered and avoided using Tolkien analogies in a more heavy-handed way,
beyond perhaps the preceding paragraph.
For clarity I instead retained the original `keyring` analogy (see below).

## Nomenclature

In the `keyring` analogy, a __key__ stores one username-password combination.
You might need more than one key per system,
if you use more than one username to access it.
A __keyring__ is a store for one or more keys.
So, conceivably, you could have multiple personas on your machine,
each with their own keyring, each accessing multiple systems.

`keyring` allows other possibilities as well.
You could have multiple keys for a given system, all on the same keyring.
Or you could use no keyring at all;
this would allow you to keep any number of keys, just the same,
but without a master password to lock them.

`speakfriend` requires users to use a keyring.
You can still use multiple keyrings
with `options(keyring_keyring = "your_keyring")`,
but you can use the package without specifying anything about keyrings.
`speakfriend` does not currently allow multiple usernames per system,
again as a bid for simplicity.
For any given system, you simply call `get_key("system_name")`
to get the user's one password for that sytem.

[1]: https://rstudio.com/
[2]: https://github.com/r-lib/keyring
[3]: https://twitter.com/sharon000
[4]: https://www.infoworld.com/article/3320999/r-tip-keep-your-passwords-and-tokens-secure-with-the-keyring-package.html
[5]: https://twitter.com/EmilyRiederer
[6]: https://emilyriederer.netlify.com/post/resource-roundup-r-in-industry-edition/
[7]: https://CRAN.R-project.org
[8]: https://github.com/BenjaminWolfe/speakfriend
[9]: http://ae-lib.org.ua/texts-c/tolkien__the_lord_of_the_rings_1__en.htm
[10]: http://tolkiengateway.net/wiki/Doors_of_Durin#History
